Address Clustering Heuristics for Account-Based Blockchain Networks: An Analysis based on a Decentraland User Set

### 2023-11-07 paper.tex
Update section 5-7

\section{Data Analysis}
This section analyzes clustering heuristics applicable to our address set using the collected data. These include self-authorization, node embedding, and our proposed approach involving LAND transfers.

\subsection{Self-Authorization}
I analyze the addresses' transaction data to find self-authorizations, focusing on the transaction's ''calldata''. The first four bytes of the calldata represent the function selector, which unambiguously specifies the smart contract function the transaction originator wants to invoke. \newline
\textit{First}, I identify all transactions that begin with a function signature related to the approval functions corresponding to ERC-20, ERC-721, and ERC-1155 token standards. \textit{Second}, I extract the approved (spender) address, which is always located at the same position within the calldata due to token standardization. I then record the spender address in a new field within our data frame. \textit{Third}, I implement two filters on the approval transactions. The first filter checks if the spender address is part of our address set, and the second ensures that the spender address is different from the address initiating the transaction, which may occur due to a user interface bug or testing transactions. \textit{Fourth}, I group the remaining approvals by the initiating address and form the clusters.

@@

\subsection{Node Embeddings}
Our analysis of node embedding methods starts with a network graph built from intra-set asset transfers. I treat addresses as nodes and add undirected edges between them if they interacted with each other. \newline
To use the Python library developed by \cite{karateclub}, I remove multi-edges and loops (i.e., transfers with the same sender and recipient address) and only consider nodes within the largest connected component. The resulting network graph includes 51,566 nodes and 249,302 edges. As previously stated, I utilize the Diff2Vec, Role2Vec, and DeepWalk methods from the library. \newline
Subsequently, I ran ten independent experiments for each method. Our choice of input parameters is identical to \cite{Beres2020}. Each experiment returns a node embedding in a matrix with 51,566 rows corresponding to the number of nodes in the graph and 128 columns representing the feature vector dimensions. \newline
For any node within the embedding, I can easily order all other nodes by the Euclidean distance. I achieve this through the FAISS Python library, which is suitable for efficient similarity search and clustering of dense vectors \citep{johnson2019faiss}.\newline 
This feature enables us to search for our ENS ground truth pairs and report their relative rank. The results for each address pair, averaged over ten experiments, are presented in Table \ref{tbl:ENS_Domain_Ranks}. Additionally, I calculated the mean, median, standard deviation, and the rate of the target address appearing within the five or ten nearest neighbors for each method.

All three methods demonstrated remarkable efficacy in locating the target address, which indicates their ability to find multiple addresses controlled by a single entity.
Role2Vec is the most effective method, with the target address, on average, found within the 28 nearest neighbors. In contrast, Diff2Vec and DeepWalk produced slightly higher averages of 46 and 62, respectively. Furthermore, Role2Vec found the target address within the ten nearest neighbors in 24 out of 34 cases (71\%). Diff2Vec is barely less effective than Role2Vec or DeepWalk, although it is still generally successful. \newline
Notably, two address pairs are absent within the network graph and, therefore, display no values. However, they have been included to demonstrate that the intra-set asset transfer approach may not always be applicable to cluster all addresses of a predefined set.

\begin{table}[h!]
\scriptsize
  \centering
  \begin{tabular}{lrrr}
    \hline
	\textbf{User} & \textbf{Diff2Vec} & \textbf{Role2Vec} & \textbf{DeepWalk} \\
	\hline
    1 & 2.5 & 6.2 & 2.7 \\
    2 & 577.3 & 33.3 & 58.8 \\
    3 & 2.8 & 2.5 & 3.3 \\
    4 & 106.8 & 71.1 & 20.1 \\
    5 & - & - & - \\
    6 & 11.1 & 3.0 & 1.4 \\
    7 & 30.7 & 1.0 & 1.0 \\
    8 & 44.3 & 91.2 & 545.4 \\
    9 & 16.3 & 10.4 & 4.1 \\
    10 & 1.0 & 1.0 & 1.0 \\
    11 & 129.1 & 10.0 & 605.4 \\
    12 & - & - & - \\
    13 & 1.0 & 1.0 & 1.8 \\
    14 & 1.9 & 1.0 & 1.0 \\
    15 & 10.4 & 1.0 & 18.2 \\
    16 & 4.0 & 1.0 & 10.3 \\
    17 & 4.2 & 1.5 & 27.2 \\
    18 & 5.7 & 5.6 & 108.5 \\
    19 & 52.3 & 373.8 & 9.0 \\
    20 & 9.0 & 9.0 & 51.8 \\
    21 & 474.2 & 86.7 & 2.4 \\
    22 & 1.7 & 5.0 & 1.0 \\
    23 & 2.7 & 1.0 & 2.3 \\
    24 & 1.3 & 3.4 & 2.4 \\
    25 & 1.0 & 1.0 & 2.2 \\
    26 & 3.7 & 2.0 & 284.0 \\
    27 & 16.9 & 14.4 & 2.2 \\
    28 & 3.6 & 2.6 & 1.7 \\
    29 & 1.5 & 2.0 & 1.6 \\
    30 & 2.7 & 1.0 & 1.1 \\
    31 & 2.0 & 13.6 & 101.2 \\
    32 & 64.8 & 1.1 & 1.6 \\
    33 & 3.0 & 19.3 & 1.7 \\
    34 & 11.0 & 5.6 & 3.4 \\
    35 & 52.1 & 68.9 & 47.3 \\
    36 & 10.3 & 4.0 & 227.0 \\
    \hline
    Mean & 46.038 & 28.073 & 62.084 \\
    Median & 4.2 & 3.4 & 3.3 \\
    Standard Deviation & 120.504 & 68.814 &  139.932\\
    Within 5 nearest & 17 (0.5) & 19 (0.56) & 20 (0.59) \\
    Within 10 nearest & 19 (0.56) & 24 (0.71) & 21 (0.62)\\
    \hline
  \end{tabular}
  \caption{Average Rank of Target Address}
  \label{tbl:ENS_Domain_Ranks}
\end{table}

After evaluating the different node embedding methods, we want to cluster the addresses to assess the number of entities represented by the 51,566 nodes in the network. We use the node embedding generated by Role2Vec. \newline
One approach might involve grouping a node's $k$ nearest neighbors to a cluster and subsequently merging clusters until each node is assigned to exactly one cluster. However, selecting an appropriate $k$ is challenging, and making the case that each address belongs to a cluster of at least $k$ is problematic. \newline
Instead, we choose to build clusters based on Euclidean distance, where all neighbors of a node below a certain threshold are clustered. Afterward, we merge overlapping clusters; for instance, if node A is clustered with nodes B and C, and node C is clustered with nodes A and X, we form a cluster comprising A, B, C, and X. \newline
We explore the sensitivity of clusters to various Euclidean distance threshold values. Figure \ref{fig:threshold_sensitivity} shows the histogram of cluster sizes, and Table \ref{tbl:cluster_table} displays the number of resulting clusters depending on various threshold values. We chose said values after observing the distance between ground truth pairs. For better visibility, we only display clusters below 20 addresses in Figure \ref{fig:threshold_sensitivity}. At thresholds of 0.6, 0.7, and 0.8, we find 23,608, 18,592, and 14,793 clusters, respectively. Regardless of the threshold, the largest cluster has more than 500 addresses. Unsurprisingly, higher thresholds result in fewer single-address clusters and more large-sized clusters. The effect of the threshold is most pronounced for single address clusters while affecting other cluster sizes much less.

@@

\subsection{LAND Transfers}
To implement the LAND transfer clustering heuristic, we select all transaction hashes of LAND transfers that occurred between two Decentraland addresses. Next, we gather all token and native asset transfers associated with the selected transaction hashes. We group the data by transaction hash and examine each transfer's sender (\texttt{from}) and recipient (\texttt{to}) address. If the same address appears as the sender and recipient, we mark the transaction as a purchase. Finally, we exclude purchases and build clusters for addresses that sent each other LAND without compensation. \newline
A total of 2159 intra-set LAND transfers were recorded, of those, 677 (X \%) were without compensation. We were able to group the 413 unique addresses from these uncompensated transfers into 161 distinct clusters. 
Notably, the majority of clusters consist of two or three addresses. However, one cluster counting 71 addresses stands out. \newline 
Figure \ref{fig:LAND_network} visualizes LAND transfers between Decentraland addresses as a token network graph. In this graph, red edges signify non-compensated transfers, whereas blue edges represent LAND transfers made as part of a trade/purchase. Notably, the 71-address cluster previously mentioned is visible on the top-right of the largest connected component. \newline
A few key observations can be drawn from this network graph. In line with our expectations, most LAND transfers involve some form of compensation. Moreover, certain nodes play a central role in LAND trading activity, indicated by their high connectivity within the network. Lastly, a recurring behavior pattern is observed wherein an individual address purchases LAND and subsequently transfers it to a third address.

@@

\subsection{Combining Clustering Results}
To conclude our clustering analysis, we combine the results of self-authorization, learned node embedding, and LAND transfer heuristics. By doing so, we can cluster 51,581 out of 137,544 (37.5\%) addresses representing 29,680 entities. 
For the remaining addresses, we consider each as an individual entity. 
We applied a strict distance threshold of 0.5 for clustering embeddings generated by Role2Vec and excluded the 80-address cluster identified in self-authorization.
Under these assumptions, we estimate 115,643 active users during the time frame studied. Of these, 8,483 entities (7.3\%) interacted with Decentraland through multiple addresses. The largest entity comprises 499 addresses. We plot the distribution of cluster sizes in Figure \ref{fig:barplot_clusters}.

@@

\section{Discussion}
Throughout our analysis, we successfully applied three heuristics to cluster our address set while making use of on-chain data. In the following discussion, we will examine each of the approaches step-by-step.

Implementing the \textit{self-authorization} heuristic is straightforward but results in very few identified address clusters. This is because it requires a highly specific on-chain behavior, which is rarely observed in practice. In particular, approvals are seldom used to authorize another externally owned account one might own. We also encountered a significant number of false positives due to non-standard implementations. We are very confident that the large cluster is not controlled by the same entity, and therefore remove it. After excluding these instances, we identified a total of 15 entities consisting of 35 addresses. This outcome aligns with findings by \cite{FV:17}, who also observed that the heuristic captures only a small proportion of addresses. In his analysis of Ethereum transaction data up to block number \texttt{8,500,000}, he identified 4,599 entities from 7,107 addresses.

Address clustering through \textit{graph representation learning} proved to be the most effective approach in our analysis. Unlike methods that require participation in specific on-chain events, node embedding algorithms examine the entire asset transfer network between addresses, making it broadly applicable. \newline
Our implementation of node embedding methods differs from previous approaches by \cite{Beres2020} and \cite{wu2022tutela}, the main reason being different settings and research questions. We construct the network graph based on asset transfers within our address set and simultaneously analyze Polygon and Ethereum data. \newline

@@

Moreover, we validated our approach and found better results regarding the average rank of ground truth address pairs when building the network graph from intra-set asset transfers compared to using all token transfer events and/or transactions. A possible extension would include incorporating edge weights, similar to \cite{wu2022tutela}; however, we refrained from doing this because we would have needed to implement a custom node embedding algorithm.\newline
We applied three node embedding algorithms that delivered robust results in the experiments by \cite{Beres2020} and assessed their effectiveness in detecting ground truth pairs. Our results indicate that Role2Vec is the best-performing method, locating the ground truth pair within the ten nearest neighbors in 70\% of cases.  This result diverges from the findings of \cite{Beres2020}, who determined Diff2Vec to be the superior method among the three evaluated. \newline
Determining why one approach outperforms another is challenging. DeepWalk and Diff2Vec, both neighborhood-preserving algorithms, mainly capture node proximity. As noted by \cite{ahmed2018roletovec}, this does not necessarily imply similarity, especially when addresses frequently interact without being controlled by the same entity. In contrast, Role2Vec, a structural node embedding method, identifies nodes with similar roles within the network. In our evaluation sample, network position or role may better represent node similarity and structural relatedness. Despite these insights, the differences in results could be specific to the data at hand, given the small sample size, potential evaluation metric flaws, and the overall robust performance of all methods. \newline 
Additionally, our evaluation sample may be biased, as one could argue that users changing the address associated with their ENS domain name while also keeping the reverse records are not primarily concerned with preserving privacy. \newline
We also need to consider that average rank may not be the most appropriate evaluation metric, particularly when a user controls more than two addresses. Suppose a single entity controlling 20 addresses, but only one pair of addresses is detected as ground truth. Even if the algorithm finds all 20 addresses within the nearest neighbors but identifies the ground truth pair as the last one, we will score it lower than another algorithm that finds only one of the twenty addresses within the 10 nearest neighbors. \newline
Creating clusters out of the node embeddings enables us to estimate the number of entities represented by the 51,566 addresses. Our method involves calculating the Euclidean distance between pairs of feature vectors, explicitly using the node embedding matrix generated by Role2Vec. Nodes are clustered if their distance falls below a predetermined threshold, and any overlapping clusters are subsequently merged. \newline 
While our clustering method may not be ideal, it is probably more effective than clustering a fixed number of $k$ neighbors for each node. We observed that the number of clusters and their size is highly sensitive to the chosen distance threshold, making it challenging to assess the accuracy of these clusters. We have provided a sensitivity analysis to illustrate this issue. The distances between the feature vectors tend to be narrow, with significant variation even among various ground truth pairs' distances. Regardless of the threshold value, we were able to locate a cluster of more than 500 addresses. Manual inspection revealed that these addresses show very similar on-chain patterns: They are all exclusively active on Polygon, and most transfers are mints of Decentraland-related NFTs. Further, they transfer some NFTs to address \texttt{''0xE2e2F2240a84A61B7dFF50a86ee10d5FaF7faCa8''}. The transfers to and from this address are shown in Figure \ref{fig:cluster}. Incoming transfers are indicated by green edges, while red edges show outgoing transfers. We hypothesize that the network in question may be operated by a user or even bot that creates multiple distinct addresses and actively engages with Decentraland. It seems likely that this bot may target specific locations or events where NFTs are (automatically) air-dropped to avatars in presence. If the NFTs hold value, the bot may then aggregate them to the main address and sell them on the Decentraland marketplace. This hypothesis could be verified with the Decentraland movement data. \newline

@@

Finally, we explored potential enhancements to the graph representation learning method, such as incorporating time-of-day-activity and/or gas price selection information into the feature vectors generated by Role2Vec. Unfortunately, these modifications did not lead to any significant improvements. This outcome may be attributed to challenges in implementation, particularly in scaling the values without disturbing the integrity of the dense feature vectors.  

@@

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.8\textwidth]{./figures/network_visualization_0xe2e.png}
	\caption{Token Transfer Graph for \texttt{0xE2e2F$\dots$faCa8}}
	\label{fig:cluster}
\end{figure}

The \textit{LAND transfer} heuristic is targeted at users revealing address clusters by sending high-value assets to other addresses. In this specific case, we use the information that all addresses are related to Decentraland and analyze LAND transfers. However, the approach could be applied to other NFT collections or large asset transfers in general. \newline
The main assumption behind this approach is that users may want to reorganize their accounts from time to time, primarily to increase security guarantees or spread the risk across multiple accounts. As a concrete example, some users may have started with a simple browser-based wallet, but as their assets appreciated in value, they transitioned to a more secure hardware wallet. We detect 161 entities with 413 addresses applying this approach. One cluster with 70 addresses is likely not a single entity. It includes an address that participated in the initial LAND auction and subsequently distributed LAND. This may have happened through off-chain compensation. The advantage of this method is that it is relatively easy to visualize and understand. The disadvantage is that it demands a specific on-chain behavior and is only scalable with significant effort. Additionally, many events that occur off-chain are not detectable in the data, resulting in increased false positives.

Ultimately, we combined the results for each clustering heuristic mentioned above. Consequently, we managed to cluster 51,518 addresses, corresponding to 29,680 distinct entities. When adding the unclustered addresses, we identified 115,643 active users during the nine months, with 8,483 entities interacting through multiple addresses. \newline
These results, however, are heavily influenced by the assumptions we made. We were conservative by setting a low distance threshold and treating all other addresses as individual entities. Raising the threshold would result in larger clusters and fewer entities. \newline 
The majority of addresses were clustered through graph representation learning methods. Unlike other clustering heuristics, this approach does not rely on specific on-chain behavior patterns or participation in particular events. Our implementation generally captures all addresses that transferred assets within the address set at least once. Nevertheless, some clusters may still go undetected, while others may be overestimated. It is important to note that these methods are heuristic, and testing their efficacy on a large scale is not feasible. \newline
Role2Vec performed best for our small test sample but may not always superior to Diff2Vec or DeepWalk. Incorporating time-of-day and normalized gas into feature vectors or refining clusters using a two-step approach requires further research but could enhance clustering results.

>>>>

\section{Data Analysis}
This section analyzes clustering heuristics applicable to the address set using the collected data. These include self-authorization, node embedding, and the proposed approach involving LAND transfers.

\subsection{Self-Authorization}
I analyze the addresses' transaction data to find self-authorizations, focusing on the transaction's ''calldata''. The first four bytes of the calldata represent the function selector, which unambiguously specifies the smart contract function the transaction originator wants to invoke. \newline
\textit{First}, I identify all transactions that begin with a function signature related to the approval functions corresponding to ERC-20, ERC-721, and ERC-1155 token standards. \textit{Second}, I extract the approved (spender) address, which is always located at the same position within the calldata due to token standardization. The spender address is recorded in a new field within the data frame. \textit{Third}, I implement two filters on the approval transactions. The first filter checks if the spender address is part of the address set, and the second ensures that the spender address is different from the address initiating the transaction, which may occur due to a user interface bug or testing transactions. \textit{Fourth}, I group the remaining approvals by the initiating address and form the clusters.

@@

\subsection{Node Embeddings}
The analysis of node embedding methods starts with a network graph built from intra-set asset transfers. Addresses are treated as nodes and undirected edges are added if they interacted with each other. \newline
To use the Python library developed by \cite{karateclub}, I remove multi-edges and loops (i.e., transfers with the same sender and recipient address) and only consider nodes within the largest connected component. The resulting network graph includes 51,566 nodes and 249,302 edges. As previously stated, I utilize the Diff2Vec, Role2Vec, and DeepWalk methods from the library. \newline
Subsequently, I ran ten independent experiments for each method. The choice of input parameters is identical to \cite{Beres2020}. Each experiment returns a node embedding in a matrix with 51,566 rows corresponding to the number of nodes in the graph and 128 columns representing the feature vector dimensions. \newline
For any node within the embedding, I can easily order all other nodes by the Euclidean distance. I achieve this through the FAISS Python library, which is suitable for efficient similarity search and clustering of dense vectors \citep{johnson2019faiss}.\newline 
This enables searching for the ENS ground truth pairs and reporting their relative rank. The results for each address pair, averaged over ten experiments, are presented in Table \ref{tbl:ENS_Domain_Ranks}. Additionally, I calculated the mean, median, standard deviation, and the rate of the target address appearing within the five or ten nearest neighbors for each method.

All three methods demonstrated remarkable efficacy in locating the target address, which indicates their ability to find multiple addresses controlled by a single entity.
Role2Vec is the most effective method, with the target address, on average, found within the 28 nearest neighbors. In contrast, Diff2Vec and DeepWalk produced slightly higher averages of 46 and 62, respectively. Furthermore, Role2Vec found the target address within the ten nearest neighbors in 24 out of 34 cases (71\%). Diff2Vec is barely less effective than Role2Vec or DeepWalk, although it is still generally successful. \newline

\begin{table}[h!]
\scriptsize
  \centering
  \begin{tabular}{lrrr}
    \hline
	\textbf{User} & \textbf{Diff2Vec} & \textbf{Role2Vec} & \textbf{DeepWalk} \\
	\hline
    1 & 2.5 & 6.2 & 2.7 \\
    2 & 577.3 & 33.3 & 58.8 \\
    3 & 2.8 & 2.5 & 3.3 \\
    4 & 106.8 & 71.1 & 20.1 \\
    5 & - & - & - \\
    6 & 11.1 & 3.0 & 1.4 \\
    7 & 30.7 & 1.0 & 1.0 \\
    8 & 44.3 & 91.2 & 545.4 \\
    9 & 16.3 & 10.4 & 4.1 \\
    10 & 1.0 & 1.0 & 1.0 \\
    11 & 129.1 & 10.0 & 605.4 \\
    12 & - & - & - \\
    13 & 1.0 & 1.0 & 1.8 \\
    14 & 1.9 & 1.0 & 1.0 \\
    15 & 10.4 & 1.0 & 18.2 \\
    16 & 4.0 & 1.0 & 10.3 \\
    17 & 4.2 & 1.5 & 27.2 \\
    18 & 5.7 & 5.6 & 108.5 \\
    19 & 52.3 & 373.8 & 9.0 \\
    20 & 9.0 & 9.0 & 51.8 \\
    21 & 474.2 & 86.7 & 2.4 \\
    22 & 1.7 & 5.0 & 1.0 \\
    23 & 2.7 & 1.0 & 2.3 \\
    24 & 1.3 & 3.4 & 2.4 \\
    25 & 1.0 & 1.0 & 2.2 \\
    26 & 3.7 & 2.0 & 284.0 \\
    27 & 16.9 & 14.4 & 2.2 \\
    28 & 3.6 & 2.6 & 1.7 \\
    29 & 1.5 & 2.0 & 1.6 \\
    30 & 2.7 & 1.0 & 1.1 \\
    31 & 2.0 & 13.6 & 101.2 \\
    32 & 64.8 & 1.1 & 1.6 \\
    33 & 3.0 & 19.3 & 1.7 \\
    34 & 11.0 & 5.6 & 3.4 \\
    35 & 52.1 & 68.9 & 47.3 \\
    36 & 10.3 & 4.0 & 227.0 \\
    \hline
    Mean & 46.038 & 28.073 & 62.084 \\
    Median & 4.2 & 3.4 & 3.3 \\
    Standard Deviation & 120.504 & 68.814 &  139.932\\
    Within 5 nearest & 17 (0.5) & 19 (0.56) & 20 (0.59) \\
    Within 10 nearest & 19 (0.56) & 24 (0.71) & 21 (0.62)\\
    \hline
  \end{tabular}
  \caption{Average Rank of Target Address. Two address pairs are not part of the largest connected component within the network graph and therefore display no values.}
  \label{tbl:ENS_Domain_Ranks}
\end{table}

After evaluating the different node embedding methods, I cluster the addresses to assess the number of entities represented by the 51,566 nodes in the network. The node embeddings generated by Role2Vec are used. \newline
One approach might involve grouping a node's $k$ nearest neighbors to a cluster and subsequently merging clusters until each node is assigned to exactly one cluster. However, selecting an appropriate $k$ is challenging, and making the case that each address belongs to a cluster of at least $k$ is problematic. \newline
Instead, I choose to build clusters based on Euclidean distance, where all neighbors of a node below a certain threshold are clustered. Afterward, the overlapping clusters are merged; for instance, if node A is clustered with nodes B and C, and node C is clustered with nodes A and X, a cluster comprising A, B, C, and X is formed. \newline
I explore the sensitivity of clusters to various Euclidean distance threshold values. Figure \ref{fig:threshold_sensitivity} shows the histogram of cluster sizes, and Table \ref{tbl:cluster_table} displays the number of resulting clusters depending on various threshold values. I chose said values after observing the distance between ground truth pairs. For better visibility, only clusters below 20 addresses are displayed in Figure \ref{fig:threshold_sensitivity}. At thresholds of 0.6, 0.7, and 0.8, I find 23,608, 18,592, and 14,793 clusters, respectively. Regardless of the threshold, the largest cluster has more than 500 addresses. By design, higher thresholds result in fewer single-address clusters and more large-sized clusters. The effect of the threshold is most pronounced for single address clusters while affecting other cluster sizes much less.

@@

\subsection{LAND Transfers}
To implement the LAND transfer clustering heuristic, I select all transaction hashes of LAND transfers that occurred between two Decentraland addresses. Next, I gather all token and native asset transfers associated with the selected transaction hashes. The data is grouped by transaction hash and each transfer's sender (\texttt{from}) and recipient (\texttt{to}) address is examined. If the same address appears as the sender and recipient, I mark the transaction as a purchase. Finally, purchases are excluded and clusters for addresses that sent each other LAND without compensation are built. \newline
A total of 2159 intra-set LAND transfers were recorded, of those, 677 (X \%) were without compensation. We were able to group the 413 unique addresses from these uncompensated transfers into 161 distinct clusters. 
Notably, the majority of clusters consist of two or three addresses. However, one cluster counting 71 addresses stands out. \newline 
Figure \ref{fig:LAND_network} visualizes LAND transfers between Decentraland addresses as a token network graph. In this graph, red edges signify non-compensated transfers, whereas blue edges represent LAND transfers made as part of a trade/purchase. Notably, the 71-address cluster previously mentioned is visible on the top-right of the largest connected component. \newline
A few key observations can be drawn from this network graph. In line with our expectations, most LAND transfers involve some form of compensation. Moreover, certain nodes play a central role in LAND trading activity, indicated by their high connectivity within the network. Lastly, a recurring behavior pattern is observed wherein an individual address purchases LAND and subsequently transfers it to a third address.

@@

\subsection{Combining Clustering Results}
To conclude the clustering analysis, the results of self-authorization, learned node embedding, and LAND transfer heuristics are combined. By doing so, I can cluster 51,581 out of 137,544 (37.5\%) addresses representing 29,680 entities. 
For the remaining addresses, I consider each as an individual entity. 
I applied a strict distance threshold of 0.5 for clustering embeddings generated by Role2Vec and excluded the 80-address cluster identified in self-authorization.
Under these assumptions, I estimate 115,643 active users during the time frame studied. Of these, 8,483 entities (7.3\%) interacted with Decentraland through multiple addresses. The largest entity comprises 499 addresses. The distribution of the cluster sizes for different thresholds is shown in Figure \ref{fig:barplot_clusters}.

@@

\section{Discussion}
Throughout the analysis, I successfully applied three heuristics to cluster our address set while making use of on-chain data. In the following discussion, I will examine each of the approaches step-by-step.

Implementing the \textit{self-authorization} heuristic is straightforward but results in very few identified address clusters. This heuristic demands a highly specific on-chain behavior, which is rarely observed in practice. In particular, approvals are seldom used to authorize another externally owned account one might own. I also encountered a significant number of false positives due to non-standard implementations. I am very confident that the large cluster is not controlled by the same entity, and therefore remove it. After excluding these instances, I identified a total of 15 entities consisting of 35 addresses. This outcome aligns with findings by \cite{FV:17}, who also observed that the heuristic captures only a small proportion of addresses. In his analysis of Ethereum transaction data up to block number \texttt{8,500,000}, he identified 4,599 entities from 7,107 addresses.

Address clustering through \textit{graph representation learning} proved to be the most effective approach in my analysis. Unlike methods that require participation in specific on-chain events, node embedding algorithms examine the entire asset transfer network between addresses, making it broadly applicable. \newline
The implementation of the node embedding methods differs from previous approaches by \cite{Beres2020} and \cite{wu2022tutela}, the main reason being different settings and research questions. I construct the network graph based on asset transfers within our address set and simultaneously analyze Polygon and Ethereum data. \newline

@@

Moreover, I validated the approach and found better results regarding the average rank of ground truth address pairs when building the network graph from intra-set asset transfers compared to using all token transfer events and/or transactions. A possible extension would include incorporating edge weights, similar to \cite{wu2022tutela}; however, I refrained from doing this because the implementation a custom node embedding algorithm would have been necessary.\newline
I applied three node embedding algorithms that delivered robust results in the experiments by \cite{Beres2020} and assessed their effectiveness in detecting ground truth pairs. The results indicate that Role2Vec is the best-performing method, locating the ground truth pair within the ten nearest neighbors in 70\% of cases.  This result diverges from the findings of \cite{Beres2020}, who determined Diff2Vec to be the superior method among the three evaluated. \newline
Determining why one approach outperforms another is challenging. DeepWalk and Diff2Vec, both neighborhood-preserving algorithms, mainly capture node proximity. As noted by \cite{ahmed2018roletovec}, this does not necessarily imply similarity, especially when addresses frequently interact without being controlled by the same entity. In contrast, Role2Vec, a structural node embedding method, identifies nodes with similar roles within the network. In my evaluation sample, network position or role may better represent node similarity and structural relatedness. Despite these insights, the differences in results could be specific to the data at hand, given the small sample size, potential evaluation metric flaws, and the overall robust performance of all methods. \newline 
Additionally, the evaluation sample may be biased, as one could argue that users changing the address associated with their ENS domain name while also keeping the reverse records are not primarily concerned with preserving privacy. \newline
It should also be noted that average rank may not be the most appropriate evaluation metric, particularly when a user controls more than two addresses. Suppose a single entity controlling 20 addresses, but only one pair of addresses is detected as ground truth. Even if the algorithm finds all 20 addresses within the nearest neighbors but identifies the ground truth pair as the last one, we will score it lower than another algorithm that finds only one of the twenty addresses within the 10 nearest neighbors. \newline
Creating clusters out of the node embeddings enables an estimation of the number of entities represented by the 51,566 addresses. The method involves calculating the Euclidean distance between pairs of feature vectors, explicitly using the node embedding matrix generated by Role2Vec. Nodes are clustered if their distance falls below a predetermined threshold, and any overlapping clusters are subsequently merged. \newline 
While the clustering method may not be ideal, it is probably more effective than clustering a fixed number of $k$ neighbors for each node. I observed that the number of clusters and their size is highly sensitive to the chosen distance threshold, making it challenging to assess the accuracy of these clusters. I have provided a sensitivity analysis to illustrate this issue. The distances between the feature vectors tend to be narrow, with significant variation even among various ground truth pairs' distances. Regardless of the threshold value, a cluster of more than 500 addresses persisted. Manual inspection revealed that these addresses show very similar on-chain patterns: They are all exclusively active on Polygon, and most transfers are mints of Decentraland-related NFTs. Further, they transfer some NFTs to address \texttt{''0xE2e2F224\dots''}. The transfers to and from this address are shown in Figure \ref{fig:cluster}. Incoming transfers are indicated by green edges, while red edges show outgoing transfers. I hypothesize that the network in question may be operated by a user or even bot that creates multiple distinct addresses and actively engages with Decentraland. It seems likely that this user may target specific locations or events where NFTs can be collected or are (automatically) air-dropped to avatars in presence. If the NFTs hold value, the user may then aggregate them to the main address and sell them on the Decentraland marketplace. This hypothesis could be verified using Decentraland movement data. \newline

@@

Finally, I explored potential enhancements to the graph representation learning method, such as incorporating time-of-day-activity and/or gas price selection information into the feature vectors generated by Role2Vec. Unfortunately, these modifications did not lead to any significant improvements. This outcome may be attributed to challenges in implementation, particularly in scaling the values without disturbing the integrity of the dense feature vectors.  

@@

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.8\textwidth]{./figures/network_visualization_0xe2e.png}
	\caption{Token Transfer Graph for \texttt{0xE2e2F24$\dots$faCa8}}
	\label{fig:cluster}
\end{figure}

The \textit{LAND transfer} heuristic is targeted at users revealing address clusters by sending high-value assets to other addresses. In this specific case, I use the information that all addresses are related to Decentraland and analyze LAND transfers. However, the approach could be applied to other NFT collections or large asset transfers in general. \newline
The main assumption behind this approach is that users may want to reorganize their accounts from time to time, primarily to increase security guarantees or spread the risk across multiple accounts. As a concrete example, some users may have started with a simple browser-based wallet, but as their assets appreciated in value, they transitioned to a more secure hardware wallet. We detect 161 entities with 413 addresses applying this approach. One cluster with 70 addresses is likely not a single entity. It includes an address that participated in the initial LAND auction and subsequently distributed LAND. This may have happened through off-chain compensation. The advantage of this method is that it is relatively easy to visualize and understand. The disadvantage is that it demands a specific on-chain behavior and is only scalable with significant effort. Additionally, many events that occur off-chain are not detectable in the data, resulting in increased false positives.

Ultimately, I combined the results for each clustering heuristic mentioned above. Consequently, I managed to cluster 51,518 addresses, corresponding to 29,680 distinct entities. When adding the unclustered addresses, 115,643 active users were identified during the nine months, with 8,483 entities interacting through multiple addresses. \newline
These results, however, are heavily influenced by the assumptions made. I was conservative by setting a low distance threshold and treating all other addresses as individual entities. Raising the threshold would result in larger clusters and fewer entities. \newline 
The majority of addresses were clustered through graph representation learning methods. Unlike other clustering heuristics, this approach does not rely on specific on-chain behavior patterns or participation in particular events. My implementation generally captures all addresses that transferred assets within the address set at least once. Nevertheless, some clusters may still go undetected, while others may be overestimated. It is important to note that these methods are heuristic, and testing their efficacy on a large scale is not feasible. \newline
Role2Vec performed best for our small test sample but may not always superior to Diff2Vec or DeepWalk. Incorporating time-of-day and normalized gas into feature vectors or refining clusters using a two-step approach requires further research but could enhance clustering results.
